---
title: "Correlation and Multiple Linear Regression Assignment"
author: "Laura Parsons"
date: "May 26, 2019"
output: word_document
editor_options: 
  chunk_output_type: console
---
Load the necessary libraries.  
```{r}
#install.packages("MASS")
#install.packages("leaps")
 library(tidyverse, quietly = TRUE)
library(GGally)
library(MASS)
library(leaps)
```

Task 1: Read in the data and convert season to a factor and rename the levels.
```{r}
bike= read_csv("hour.csv")
bike = bike %>% 
  mutate(season = as_factor(as.character(season))) %>%
mutate(season = fct_recode(season,
"Spring" = "1",
"Summer" = "2",
"Fall" = "3",
"Winter" = "4"))
```

Task 1 continued:Convert yr, mnth, and hr to factors. Convert and recode holiday, workingday, weathersit, and weekday.
```{r}
bike = bike %>% 
  mutate(yr = as_factor(as.character(yr))) %>%
  mutate(mnth = as_factor(as.character(mnth))) %>%
  mutate(hr = as_factor(as.character(hr))) %>%
  mutate(holiday = as_factor(as.character(holiday))) %>%
mutate(holiday = fct_recode(holiday,
"NotHoliday" = "0",
"Holiday" = "1")) %>%
  mutate(workingday = as_factor(as.character(workingday))) %>%
mutate(workingday = fct_recode(workingday,
"NotWorkingDay" = "0",
"WorkingDay" = "1")) %>%
  mutate(weathersit = as_factor(as.character(weathersit))) %>%
mutate(weathersit = fct_recode(weathersit,
"NoPrecip" = "1",
"Misty" = "2",
"LightPrecip" = "3",
"HeavyPrecip" = "4")) %>%
  mutate(weekday = as_factor(as.character(weekday))) %>%
mutate(weekday = fct_recode(weekday,
"Sunday" = "0",
"Monday" = "1",
"Tuesday" = "2",
"Wednesday" = "3",
"Thursday" = "4",
"Friday" = "5",
"Saturday" = "6"))
```

We converted yr, mnth, and hr to factors instead of leaving them as numbers because they are categorical variables. Even though these variables contain numerical values since you would not use year, month or hour within a calculation they are then considered qualitative variables.  


Task 2:Which quantitative variables appear to be best correlated with count?
```{r}
str(bike)
ggpairs(bike, columns = c(1,11:14,17))
```

The quantitative variables that appear to be best correlated with count are temp and atemp. This is followed by hum and instant. 

Task 3:Complete a box plot for each categorical variable, Which of these affect count. 
```{r}
ggplot(bike,aes(x=hr,y=count)) + geom_boxplot()
```

In the box plot for hour versus count we can see that hour does affect count and should probably be used in a regression model. Hour affects count because the count increases at key points of the day for example between 7am and 9 am when people wake up and may be traveling to work. There also seems to be a small peak around lunch time (11 and 12) as well as a peak at the end of the work day (starting at 16, continuing at 17 and decreasing at 18).  

```{r}
ggplot(bike,aes(x=season,y=count)) + geom_boxplot()
```

In the box plot for season versus count we can see that season does affect count and should probably be used in a regression model. Season affects count because the count increases at key seasons throughout the year. As you can see the average amount of bike rides increases in fall and summer and is lower in winter and spring. 

```{r}
ggplot(bike,aes(x=yr,y=count)) + geom_boxplot()
```

In the box plot for year versus count we can see that year does affect count and should probably be used in a regression model. Year affects count because the count is significantly higher in year 1 than in year 0. Year might affect count because of current popular health trends within the different years.  

```{r}
ggplot(bike,aes(x=mnth,y=count)) + geom_boxplot()
```

In the box plot for month versus count we can see that month does affect count and should probably be used in a regression model. Month affects count because the count increases at key months throughout the year. As you can see the average amount of bike rides increases in months 5-9. 

```{r}
ggplot(bike,aes(x=holiday,y=count)) + geom_boxplot()
```

In the box plot for holiday versus count it does not appear that holiday has a strong affect on count. Even though holiday appears to only slightly affect count it should probably still be used in a regression model. Holiday slightly affects count because the count is higher on days that are not holidays than on holidays. This indicates that the amount of bike rides is slightly related to people biking to and from work. 

```{r}
ggplot(bike,aes(x=weekday,y=count)) + geom_boxplot()
```

In the box plot for weekday versus count it does not appear that weekday has a strong affect on count. Even though weekday appears to only slightly affect count it should probably still be used in a regression model. Weekday slightly affects count because the count is higher on days that are in the average work week (Monday-Friday) and slightly lower on Saturday and Sunday. This indicates that the amount of bike rides is slightly related to people biking to and from work.  

```{r}
ggplot(bike,aes(x=workingday,y=count)) + geom_boxplot()
```

In the box plot for workingday versus count it does not appear that workingday has a strong affect on count. Even though workingday appears to only slightly affect count it should probably still be used in a regression model unless it has a strong correlation with another variable in the model. I believe that workingday will be highly correlated with variables such as weekday and holiday and that we should test for multicollinearity. Workingday slightly affects count because the count is higher on days that are working days than days that are not working days. This indicates that the amount of bike rides is slightly related to people biking to and from work.  

```{r}
ggplot(bike,aes(x=weathersit,y=count)) + geom_boxplot()
```

In the box plot for weathersit versus count it does appear that weathersit has a strong affect on count and should probably be used in a regression model. Weathersit affects count because the count is significantly higher on days with little to no precipitation than on days with light and heavy precipitation. Weathersit most likely affects the count of bike rides because bike riding is an outdoor activity and people prefer not to be outdoors during inclement weather. 

Task 4: Use forward stepwise regression to build a multiple linear regression model to predict count.
```{r}
bike2=bike  %>%
  dplyr::select(-c(instant, dteday, registered, casual))
allmod = lm(count ~., bike2)
summary(allmod)
emptymod = lm(count ~1, bike2) 
summary(emptymod)
forwardmod = stepAIC(emptymod, direction = "forward", scope=list(upper=allmod,lower=emptymod),
                      trace = TRUE)  
summary(forwardmod)
```

The forward model includes the variables hr, atemp, yr, weathersit, season, month, hum, weekday, holiday, windspeed and temp. This seems to be a good quality model however, it did include holiday and weekday which although did have an affect on count seemed to have a small affect on count. This appears to be a high quality model most of the variables appear to be significant (if not that is because the other values within that variable are significant). It is odd that most of the weekdays have negative coefficients considering we would expect that certain days of the work week (Monday, Tuesday, Wednesday, and Thursday) to be positive. This indicates that there may be some slight correlation between the variables. Overall this appears to be a good model and appears to match our intuition. 

Task 5: Use backward stepwise regression to build a multiple linear regression model to predict count. 
```{r}
bike2=bike  %>%
  dplyr::select(-c(instant, dteday, registered, casual))
allmod = lm(count ~., bike2)
summary(allmod)
emptymod = lm(count ~1, bike2) 
summary(emptymod)
backmod = stepAIC(allmod, direction = "backward", trace = TRUE) 
summary(backmod)
```

The backward model includes the variables season, yr, mnth, hr, holiday, weekday, weathersit, temp, atemp, hum, and windspeed. This model is the same as the forward model.   

The variable workingday appears to be perfectly correlated with another variable and is being kicked out of the model. However, workingday is represented in the model through the variables weekday and holiday. 

Task 6:Convert Year into an integer and rerun forward stepwise regression. 
```{r}
bike = bike %>% mutate(yr = as.integer(yr)-1)
bike2=bike  %>%
  dplyr::select(-c(instant, dteday, registered, casual))
allmod = lm(count ~., bike2)
summary(allmod)
emptymod = lm(count ~1, bike2) 
summary(emptymod)
forwardmod = stepAIC(emptymod, direction = "forward", scope=list(upper=allmod,lower=emptymod),
                      trace = TRUE)  
summary(forwardmod)
```

After converting year into an integer and re-running the forward model we see that the model did not change. The forward stepwise regression model includes the variables hr, atemp, yr, weathersit, season, month, hum, weekday, holiday, windspeed and temp. 