---
title: "Project Part 1"
author: "Laura Boggs"
date: "June 17, 2019"
output: word_document
editor_options: 
  chunk_output_type: console
---
Load in the packages and the dataset.
```{r}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(GGally)#ggpairs
library(mice) #package for imputation
library(VIM)
rain= read_csv("rain.csv")
```

Observe the data.
```{r}
glimpse(rain)
summary(rain)
str(rain)
```

Observe the missingness.
```{r}
vim_plot = aggr(rain, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```

Column-wise deletion of the "cloud9am and cloud3pm" variables.  
```{r}
rain = rain %>% 
  select(-Cloud9am) %>%
  select(-Cloud3pm)
#summary(rain)
vim_plot = aggr(rain, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```

Row-wise deletion of the rows with missing values for the categorical variables.
```{r}
rain = rain %>% 
  drop_na(WindGustDir)  %>% 
  drop_na(WindDir9am) %>%
  drop_na(WindDir3pm) %>%
  drop_na(RainToday)
vim_plot = aggr(rain, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```

Use the "mice" package to impute the missing quantitative variables.  
```{r}
rain_imp = mice(rain, m=1, method = "pmm", seed=12345)
summary(rain_imp)
densityplot(rain_imp)
rain_complete = complete(rain_imp)
#summary(rain_complete)
#str(rain_complete)
```

Since the density plots for Temperature 9am and 3pm do not have a similar size and shape it leads us to believe that the imputed values are unrealistic. So I then used row wise deletion for Temperature 9am and 3pm. I then re-impute the other the missing quantitative variables. 
```{r}
rain = rain %>% 
  drop_na(Temp9am)  %>% 
  drop_na(Temp3pm)
vim_plot = aggr(rain, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)

rain_imp = mice(rain, m=1, method = "pmm", seed=12345)
summary(rain_imp)
densityplot(rain_imp)
rain_complete = complete(rain_imp)
```

Visualize the relationships between the variables and Rain Tomorrow. 
```{r}
ggplot(rain_complete, aes(x=RainTomorrow, y= MinTemp)) + geom_boxplot() + theme_bw()
ggplot(rain_complete, aes(x=RainTomorrow, y= MaxTemp)) + geom_boxplot() + theme_bw()
ggplot(rain_complete, aes(x=RainTomorrow, y= Rainfall)) + geom_boxplot() + theme_bw()+scale_y_continuous(limits = c(0, 25))

ggplot(rain_complete, aes(x=WindGustDir, fill = RainTomorrow)) + geom_bar(position="fill") + theme_bw()
t1 = table(rain_complete$RainTomorrow, rain_complete$WindGustDir) 
prop.table(t1, margin = 2 )
ggplot(rain_complete, aes(x=WindDir9am, fill = RainTomorrow)) + geom_bar(position="fill") + theme_bw()
t1 = table(rain_complete$RainTomorrow, rain_complete$WindDir9am) 
prop.table(t1, margin = 2 )
ggplot(rain_complete, aes(x=WindDir3pm, fill = RainTomorrow)) + geom_bar(position="fill") + theme_bw()
t1 = table(rain_complete$RainTomorrow, rain_complete$WindDir3pm) 
prop.table(t1, margin = 2 )

ggplot(rain_complete, aes(x=RainTomorrow, y= WindGustSpeed)) + geom_boxplot() + theme_bw()
ggplot(rain_complete, aes(x=RainTomorrow, y= WindSpeed9am)) + geom_boxplot() + theme_bw()
ggplot(rain_complete, aes(x=RainTomorrow, y= WindSpeed3pm)) + geom_boxplot() + theme_bw()

ggplot(rain_complete, aes(x=RainTomorrow, y= Humidity9am)) + geom_boxplot() + theme_bw()
ggplot(rain_complete, aes(x=RainTomorrow, y= Humidity3pm)) + geom_boxplot() + theme_bw()
ggsave("humidity3pm.png")


ggplot(rain_complete, aes(x=RainTomorrow, y= Pressure9am)) + geom_boxplot() + theme_bw()
ggplot(rain_complete, aes(x=RainTomorrow, y= Pressure3pm)) + geom_boxplot() + theme_bw()

ggplot(rain_complete, aes(x=RainTomorrow, y= Temp9am)) + geom_boxplot() + theme_bw()
ggplot(rain_complete, aes(x=RainTomorrow, y= Temp3pm)) + geom_boxplot() + theme_bw()

ggplot(rain_complete, aes(x=RainToday, fill = RainTomorrow)) + geom_bar(position="fill") + theme_bw()
t1 = table(rain_complete$RainTomorrow, rain_complete$RainToday) 
prop.table(t1, margin = 2 )
ggsave("raintoday.png")
```
 
